### 2023-02-21 figures/fig-heatmap_identified.tex

%% Heatmap of the identified relationships between perceptions & conversation architecture

\begin{figure*}[]
  \centering
  \includegraphics[width=\textwidth]{figures/fig-heatmap_identified.pdf}
  \caption{Heatmap of literature for perception of conversational agents for each type of conversational architecture element}
  \label{fig:heatmap_identified}
\end{figure*}