### 2023-02-21 figures/fig-prisma.tex

%% PRISMA Process Visualization %%

\begin{figure}[h]
  \centering
  \includegraphics[width=1\columnwidth]{figures/fig-prisma.pdf}
  \caption{Literature search using PRISMA guidelines}
  \label{fig:prisma}
\end{figure}