### 2023-02-21 sections/6-conclusion.tex

%% CONCLUSION SECTION %%

\section{Conclusion}

In this paper, we discussed the results of a systematic review of existing literature published in the ACM Digital Library on the effect of conversation architecture elements on the perceptions of CAs. Through our synthesis of 57 papers in our corpus, we present a framework outlining the relationships between elements of agents’ conversation architecture (dialog strategy, content affectiveness, content style and speech format) and aspects of users’ perception (interaction, ability, sociability and humanness). During our analysis, we found the need for consistent protocols in evaluating perceptions of agents, as measurements are inconsistent across studies. Further research is needed to investigate the under-explored areas in the framework, the relationship across perception aspects, the influence of contextual factors, and the effect of composite conversation architecture elements on users' perceptions. While our research contribute to the design conversation architecture to orchestrate specific perceptions of agent, we urge designers to incorporate ethical perspectives into their design considerations.