### 2023-02-23 figures/fig-heatmap_coverage.tex
update introduction & related work; descriptions for figures

%% Side by side comparison between coverage and identified relationships

\begin{figure*}[]
  \centering
  \includegraphics[width=\textwidth]{figures/fig-heatmap_coverage.pdf}
  \caption{Comparison between identified relationships and explored connections in literature}
  \Description{This figure shows a side by side comparison between the 183 identified relationships and the 265 explored connections in literature for conversation architecture elements and perceptions of agents. The visuals are shown as two heatmaps, with darker shades depicting higher number of relationships. The numbers within each cell at the interaction of a specific conversation architecture element and perception aspect outlines the number of relationships found in the reviewed corpus.}
  \label{fig:heatmap_coverage}
\end{figure*}