### 2023-02-23 figures/fig-prisma.tex

%% PRISMA Process Visualization %%

\begin{figure}[h]
  \centering
  \includegraphics[width=1\columnwidth]{figures/fig-prisma.pdf}
  \caption{Literature search process}
  \Description{This figure shows the literature search process, starting with automated databases search yielding 2901 publications. Screening the titles and abstracts of the papers based selection criteria resulted in 221 papers. Next, full article reviews filtered the list down to 49 papers. An additional search identified 8 more papers, for a total of 57 papers included in our reviewed corpus.}
  \label{fig:prisma}
\end{figure}