### 2023-02-27 figures/fig-heatmap_coverage.tex
Anastasia edit

%% Side by side comparison between coverage and identified relationships

\begin{figure*}[]
  \centering
  \includegraphics[width=\textwidth]{figures/fig-heatmap_coverage.pdf}
  \caption{Number of past work that identified relationships between conversation architecture and the perception of agents (on the left) and number of past work that explored possible connections (on the right).}
  \Description{This figure shows a side-by-side comparison between the 183 identified relationships within the 265 explored connections in literature for conversation architecture elements and perceptions of agents. The visuals are shown as two heatmaps, with darker shades depicting a higher number of relationships. The numbers within each cell at the interaction of a specific conversation architecture element and perception aspect outlines the number of relationships found in the reviewed corpus.}
  \label{fig:heatmap_coverage}
\end{figure*}