### 2023-06-06 sections/7-appendix.tex

%TC:ignore
\onecolumn
\section{Appendix}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}

\begingroup
\renewcommand{\arraystretch}{1.3}

\begin{longtable}{@{} | p{0.14\textwidth} | p{0.48\textwidth} | p{0.035\textwidth} | p{0.27\textwidth} | @{}}
\caption{Articles included in literature review} \\
\hline \textbf{Authors} & \textbf{Title} & \textbf{Year} & \textbf{Venue} \\ \hline
\endfirsthead

\multicolumn{4}{c}{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\hline \textbf{Authors} & \textbf{Title} & \textbf{Year} & \textbf{Venue} \\ \hline
\endhead

\endlastfoot

%Data
\citet{andrews2012system}\cmt{[38]} & 
  System personality and persuasion in human-computer dialogue & 
  2012 & ACM Transactions on Interactive Intelligent Systems (TiiS) \\ \hline 
\citet{ashktorab2019resilient}\cmt{[88]} &
  Resilient chatbots: Repair strategy preferences for conversational breakdowns &
  2019 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{ceha2022expressive}\cmt{[77]} &
  Expressive Auditory Gestures in a Voice-Based Pedagogical Agent &
  2022 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{ceha2021can}\cmt{[57]} &
  Can a Humorous Conversational Agent Enhance Learning Experience and Outcomes? &
  2021 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{chan2021kinvoices}\cmt{[74]} &
  KinVoices: Using Voices of Friends and Family in Voice Interfaces &
  2021 & ACM Conference on Computer Supported Cooperative Work (CSCW) \\ \hline
\citet{choi2020nobody}\cmt{[54]} &
  "Nobody speaks that fast!" An empirical study of speech rate in conversational agents for people with vision impairments &
  2020 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{cox2022does}\cmt{[27]} &
  Does Chatbot Language Formality Affect Users’ Self-Disclosure? &
  2022 & ACM Conference Conversational User Interfaces (CUI) \\ \hline
\citet{cuadra2021my}\cmt{[67]} &
  My bad! repairing intelligent voice assistant errors improves interaction &
  2021 & ACM Conference on Computer Supported Cooperative Work (CSCW) \\ \hline
\citet{daher2020empathic}\cmt{[58]} &
  Empathic chatbot response for medical assistance &
  2020 & ACM International Conference on Intelligent Virtual Agents (IVA) \\ \hline
\citet{diederich2019emulating}\cmt{[25]} &
  Emulating empathetic behavior in online service encounters with sentiment-adaptive responses: insights from an experiment with a conversational agent &
  2019 & International Conference on Computers and Information Systems (ICIS) \\ \hline
\citet{dubiel2020persuasive}\cmt{[60]} &
  Persuasive synthetic speech: Voice perception and user behaviour &
  2020 & ACM Conference Conversational User Interfaces (CUI) \\ \hline
\citet{elsholz2019exploring}\cmt{[61]} &
  Exploring Language Style in Chatbots to Increase Perceived Product Value and User Engagement &
  2019 & Conference on Human Information Interaction and Retrieval (CHIIR) \\ \hline
\citet{fadhil2018effect}\cmt{[52]} &
  The effect of emojis when interacting with conversational interface assisted health coaching system &
  2018 & EAI International Conference on Pervasive Computing Technologies for Healthcare (PervasiveHealth) \\ \hline
\citet{feijoo2021effects}\cmt{[70]} &
  Effects of Virtual Humans' Gender and Spoken Accent on Users' Perceptions of Expertise in Mental Wellness Conversations &
  2021 & ACM International Conference on Intelligent Virtual Agents (IVA) \\ \hline
\citet{gnewuch2018faster}\cmt{[19]} &
  Faster is not always better: understanding the effect of dynamic response delays in human-chatbot interaction &
  2018 & European Conference on Information Systems (ECIS) \\ \hline
\citet{gnewuch2018chatbot}\cmt{[21]} &
  The Chatbot is typing…'–The Role of Typing Indicators in Human-Chatbot Interaction &
  2018 & International Conference on Computers and Information Systems (ICIS) \\ \hline
\citet{gnewuch2022opposing}\cmt{[20]} &
  Opposing Effects of Response Time in Human-Chatbot Interaction &
  2022 & Business \& Information Systems Engineering (BISE) \\ \hline
\citet{go2021conversational}\cmt{[80]} &
  Conversational Agents Replying with a Manzai-style Joke &
  2021 & Australian Conference on Human-Computer Interaction (OzCHI) \\ \hline
\citet{haas2022keep}\cmt{[78]} &
  Keep it Short: A Comparison of Voice Assistants' Response Behavior &
  2022 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{habler2019effects}\cmt{[63]} &
  Effects of Smart Virtual Assistants' Gender and Language &
  2019 & Conference on Message understanding (MuC) \\ \hline
\citet{healey2013relating}\cmt{[39]} &
  Relating to speech evoked car personalities &
  2013 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{hoegen2019end}\cmt{[31]} &
  An End-to-End Conversational Style Matching Agent &
  2019 & ACM International Conference on Intelligent Virtual Agents (IVA) \\ \hline
\citet{hu2021enhancing}\cmt{[56]} &
  Enhancing the Perceived Emotional Intelligence of Conversational Agents through Acoustic Cues &
  2021 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{hu2022polite}\cmt{[76]} &
  Polite or Direct? Conversation Design of a Smart Display for Older Adults Based on Politeness Theory &
  2022 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{huiyang2022improving}\cmt{[17]} &
  Improving Interaction Experience through Lexical Convergence: The Prosocial Effect of Lexical Alignment in Human-Human and Human-Computer Interactions &
  2022 & International Journal of Human-Computer Interaction (IJHCI) \\ \hline
\citet{jeong2019exploring}\cmt{[10]} &
  Exploring effects of conversational fillers on user perception of conversational agents &
  2019 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{jestin2022effects}\cmt{[81]} &
  Effects of Wording and Gendered Voices on Acceptability of Voice Assistants in Future Autonomous Vehicles &
  2022 & ACM Conference Conversational User Interfaces (CUI) \\ \hline
\citet{khooshabeh2011does}\cmt{[37]} &
  Does it matter if a computer jokes &
  2011 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{kim2020can}\cmt{[24]} &
  "I Can Feel Your Empathic Voice": Effects of Nonverbal Vocal Cues in Voice User Interface &
  2020 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{kim2019comparing}\cmt{[89]} &
  Comparing Data from Chatbot and Web Surveys: Effects of Platform and Conversational Style on Survey Response Quality &
  2019 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{kraus2020effects}\cmt{[64]} &
  Effects of proactive dialogue strategies on human-computer trust &
  2020 & ACM Conference on User Modeling, Adaptation and Personalization (UMAP) \\ \hline
\citet{lee2019s}\cmt{[55]} &
  What’s on Your Virtual Mind? Mind Perception in Human-Agent Negotiations &
  2019 & ACM International Conference on Intelligent Virtual Agents (IVA) \\ \hline
\citet{lee2020hear}\cmt{[23]} &
  "I Hear You, I Feel You": encouraging deep self-disclosure through a chatbot &
  2020 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{linnemann2018can}\cmt{[15]} &
  ‘Can I Trust the Spoken Dialogue System Because It Uses the Same Words as I Do?’—Influence of Lexically Aligned Spoken Dialogue Systems on Trustworthiness and User Satisfaction &
  2018 & Interacting with Computers (IWC) \\ \hline
\citet{lubis2019positive}\cmt{[43]} &
  Positive emotion elicitation in chat-based dialogue systems &
  2019 & IEEE/ACM Transactions on Audio, Speech and Language Processing (TASLP) \\ \hline
\citet{lubold2016effects}\cmt{[86]} &
  Effects of voice-adaptation and social dialogue on perceptions of a robotic learning companion &
  2016 & ACM/IEEE International Conference on Human-Robot Interaction (HRI) \\ \hline
\citet{ma2022ask}\cmt{[29]} &
  Ask a Further Question or Give a List? How Should Conversational Agents Reply to Users' Uncertain Queries &
  2022 & International Journal of Human-Computer Interaction (IJHCI) \\ \hline
\citet{miehle2018exploring}\cmt{[51]} &
  Exploring the impact of elaborateness and indirectness on user satisfaction in a spoken dialogue system &
  2018 & ACM Conference on User Modeling, Adaptation and Personalization (UMAP) \\ \hline
\citet{misu2011toward}\cmt{[83]} &
  Toward construction of spoken dialogue system that evokes users' spontaneous backchannels &
  2011 & SIGdial Workshop on Discourse and Dialogue (SIGDIAL) \\ \hline
\citet{miyamoto2017improving}\cmt{[46]} &
  Improving Relationships Based on Positive Politeness Between Humans and Life-Like Agents &
  2017 & International Conference on Human-Agent Interaction (HAI) \\ \hline
\citet{moilanen2022measuring}\cmt{[82]} &
  Measuring the Effect of Mental Health Chatbot Personality on User Engagement &
  2022 & International Conference on Mobile and Ubiquitous Multimedia (MUM) \\ \hline
\citet{niewiadomski2013laugh}\cmt{[85]} &
  Laugh-aware virtual agent and its impact on user amusement &
  2013 & International Conference on Autonomous Agents and Multiagent Systems (AAMAS) \\ \hline
\citet{ouchi2019should}\cmt{[59]} &
  Should speech dialogue system use honorific expression? &
  2019 & International Conference on Human-Agent Interaction (HAI) \\ \hline
\citet{roy2021users}\cmt{[71]} &
  Users, Tasks, and Conversational Agents: A Personality Study &
  2021 & International Conference on Human-Agent Interaction (HAI) \\ \hline
\citet{seeger2021chatbots}\cmt{[35]} &
  Chatbots often Fail! Can Anthropomorphic Design Mitigate Trust Loss in Conversational Agents for Customer Service? &
  2021 & European Conference on Information Systems (ECIS) \\ \hline
\citet{spillner2021talk}\cmt{[18]} &
  Talk to Me on My Level-Linguistic Alignment for Chatbots &
  2021 & International Conference on Human-Computer Interaction with Mobile Devices and Services (MobileHCI) \\ \hline
\citet{tolmeijer2021female}\cmt{[62]} &
  Female by default?--exploring the effect of voice assistant gender and pitch on trait and trust attribution &
  2021 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{volkel2021examining}\cmt{[69]} &
  Examining User Preference for Agreeableness in Chatbots &
  2021 & ACM Conference Conversational User Interfaces (CUI) \\ \hline
\citet{volkel2021manipulating}\cmt{[68]} &
  Manipulating and Evaluating Levels of Personality Perceptions of Voice Assistants through Enactment-Based Dialogue Design &
  2021 & ACM Conference Conversational User Interfaces (CUI) \\ \hline
\citet{volkel2022user}\cmt{[75]} &
  User perceptions of extraversion in chatbots after repeated use &
  2022 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{wester2015artificial}\cmt{[14]} &
  Artificial personality and disfluency &
  2015 & INTERSPEECH \\ \hline
\citet{westerman2019believe}\cmt{[9]} &
  I Believe in a Thing Called Bot: Perceptions of the Humanness of “Chatbots” &
  2019 & Communication Studies \\ \hline
\citet{wilhelm2022keep}\cmt{[28]} &
  Keep on Smiling: An Investigation of the Influence of the Use of Emoticons by Chatbots on User Satisfaction &
  2022 & ACM Conference Conversational User Interfaces (CUI) \\ \hline
\citet{xiao2021let}\cmt{[73]} &
  Let Me Ask You This: How Can a Voice Assistant Elicit Explicit User Feedback? &
  2021 & ACM Conference on Computer Supported Cooperative Work (CSCW) \\ \hline
\citet{yang2021effect}\cmt{[72]} &
  The Effect of the Repetitive Utterances Complexity on User’s Perceived Empathy and Desire to Continue Dialogue by a Chat-oriented Dialogue System &
  2021 & International Conference on Human-Agent Interaction (HAI) \\ \hline
\citet{yang2017perceived}\cmt{[44]} &
  Perceived Emotional Intelligence in Virtual Agents &
  2017 & ACM Conference on Human Factors in Computing Systems (CHI) \\ \hline
\citet{zhu2022effects}\cmt{[26]} &
  Effects of Emotional Expressiveness on Voice Chatbot Interactions &
  2022 & ACM Conference Conversational User Interfaces (CUI) \\ \bottomrule
%\end{tabular}
\end{longtable}
\label{tab:appendix}
\endgroup
%TC:endignore